---
title: "HateCrimeDataSubset"
author: "Obelyn Smith"
date: "2024-12-11"
output: html_document
---

```{r include=FALSE}
`table13.(1)` <- read.csv("C:/Users/obely/Downloads/table13 (1).csv", header=TRUE)
```

```{r include=FALSE}
hateCrimeData<- `table13.(1)`
```

```{r include=FALSE}
library(dplyr)
library(tidyverse)
```

```{r include=FALSE}
hateCrimeSubset<- hateCrimeData %>% filter(Agency.type=="Cities")
```

```{r include=FALSE}
hateCrimeSubset$Region<- NA
```

```{r include=FALSE}
#New England

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Connecticut"){
  hateCrimeSubset$Region[i]<- "New England"}
  
}


for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Maine"){
  hateCrimeSubset$Region[i]<- "New England"}
  
}


for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Massachusetts"){
  hateCrimeSubset$Region[i]<- "New England"}
  
}


for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="New_Hampshire"){
  hateCrimeSubset$Region[i]<- "New England"}
  
}


for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Rhode_Island"){
  hateCrimeSubset$Region[i]<- "New England"}
  
}


for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Vermont"){
  hateCrimeSubset$Region[i]<- "New England"}
  
}
```

```{r include=FALSE}
#Mideast
for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Delaware"){
  hateCrimeSubset$Region[i]<- "Mideast"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Maryland"){
  hateCrimeSubset$Region[i]<- "Mideast"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="New_Jersey"){
  hateCrimeSubset$Region[i]<- "Mideast"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="New_York"){
  hateCrimeSubset$Region[i]<- "Mideast"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Pennsylvania"){
  hateCrimeSubset$Region[i]<- "Mideast"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="DC"){
  hateCrimeSubset$Region[i]<- "Mideast"}
  
}
```

```{r include=FALSE}
#Great Lakes
for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Illinois"){
  hateCrimeSubset$Region[i]<- "Great Lakes"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Indiana"){
  hateCrimeSubset$Region[i]<- "Great Lakes"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Michigan"){
  hateCrimeSubset$Region[i]<- "Great Lakes"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Ohio"){
  hateCrimeSubset$Region[i]<- "Great Lakes"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Wisconsin"){
  hateCrimeSubset$Region[i]<- "Great Lakes"}
  
}
```

```{r include=FALSE}
#Plains
for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Iowa"){
  hateCrimeSubset$Region[i]<- "Plains"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Kansas"){
  hateCrimeSubset$Region[i]<- "Plains"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Minnesota"){
  hateCrimeSubset$Region[i]<- "Plains"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Missouri"){
  hateCrimeSubset$Region[i]<- "Plains"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Nebraska"){
  hateCrimeSubset$Region[i]<- "Plains"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="North_Dakota"){
  hateCrimeSubset$Region[i]<- "Plains"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="South_Dakota"){
  hateCrimeSubset$Region[i]<- "Plains"}
  
}
```

```{r include=FALSE}
#Southeast
for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Alabama"){
  hateCrimeSubset$Region[i]<- "Southeast"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Arkansas"){
  hateCrimeSubset$Region[i]<- "Southeast"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Florida"){
  hateCrimeSubset$Region[i]<- "Southeast"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Georgia"){
  hateCrimeSubset$Region[i]<- "Southeast"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Kentucky"){
  hateCrimeSubset$Region[i]<- "Southeast"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Louisiana"){
  hateCrimeSubset$Region[i]<- "Southeast"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Mississippi"){
  hateCrimeSubset$Region[i]<- "Southeast"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="North_Carolina"){
  hateCrimeSubset$Region[i]<- "Southeast"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="South_Carolina"){
  hateCrimeSubset$Region[i]<- "Southeast"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Tennessee"){
  hateCrimeSubset$Region[i]<- "Southeast"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Virginia"){
  hateCrimeSubset$Region[i]<- "Southeast"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="West_Virginia"){
  hateCrimeSubset$Region[i]<- "Southeast"}
  
}
```

```{r include=FALSE}
#Southwest

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Arizona"){
  hateCrimeSubset$Region[i]<- "Southwest"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="New_Mexico"){
  hateCrimeSubset$Region[i]<- "Southwest"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Oklahoma"){
  hateCrimeSubset$Region[i]<- "Southwest"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Texas"){
  hateCrimeSubset$Region[i]<- "Southwest"}
  
}
```

```{r include=FALSE}
#Rocky Mountains

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Colorado"){
  hateCrimeSubset$Region[i]<- "Rocky Mountain"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Idaho"){
  hateCrimeSubset$Region[i]<- "Rocky Mountain"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Montana"){
  hateCrimeSubset$Region[i]<- "Rocky Mountain"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Utah"){
  hateCrimeSubset$Region[i]<- "Rocky Mountain"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Wyoming"){
  hateCrimeSubset$Region[i]<- "Rocky Mountain"}
  
}
```

```{r include=FALSE}
#Far West 
for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Alaska"){
  hateCrimeSubset$Region[i]<- "Far West"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="California"){
  hateCrimeSubset$Region[i]<- "Far West"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Hawaii"){
  hateCrimeSubset$Region[i]<- "Far West"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Nevada"){
  hateCrimeSubset$Region[i]<- "Far West"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Oregon"){
  hateCrimeSubset$Region[i]<- "Far West"}
  
}

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$State[i]=="Washington"){
  hateCrimeSubset$Region[i]<- "Far West"}
  
}
```

```{r include=FALSE}
hateCrimeSubset$Pop2<- as.factor(hateCrimeSubset$Population)
levels(hateCrimeSubset$Pop2)
class(hateCrimeSubset$Pop2)
```

```{r include=FALSE}
hateCrimeSubset$Pop3 <- as.numeric(gsub(",","",hateCrimeSubset$Pop2))
class(hateCrimeSubset$Pop3)
```

```{r include=FALSE}
hateCrimeSubset$City.size<- NA
```

```{r include=FALSE}
#Very Small Towns

for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$Pop3[i]<=2500){
  hateCrimeSubset$City.size[i]<- "Very Small Towns"}
}

```

```{r include=FALSE}
#Small Towns
for (i in 1:nrow(hateCrimeSubset)) {
  if (hateCrimeSubset$Pop3[i] >= 2501 & hateCrimeSubset$Pop3[i] <= 9999) {
    hateCrimeSubset$City.size[i] <- "Small Towns"
  }
}
```

```{r include=FALSE}
#Small Cities 
for (i in 1:nrow(hateCrimeSubset)) {
  if (hateCrimeSubset$Pop3[i] >= 10000 & hateCrimeSubset$Pop3[i] <= 49999) {
    hateCrimeSubset$City.size[i] <- "Small Cities"
  }
}
```

```{r include=FALSE}
#Mid-Sized Cities 

for (i in 1:nrow(hateCrimeSubset)) {
  if (hateCrimeSubset$Pop3[i] >= 50000 & hateCrimeSubset$Pop3[i] <= 249999) {
    hateCrimeSubset$City.size[i] <- "Mid-Sized Cities"
  }
}
```

```{r include=FALSE}
for (i in 1:nrow(hateCrimeSubset)) {
  if (hateCrimeSubset$Pop3[i] >= 250000 & hateCrimeSubset$Pop3[i] <= 999999) {
    hateCrimeSubset$City.size[i] <- "Large Cities"
  }
}
```

```{r include=FALSE}
#Metropolitan Areas 
for (i in 1:nrow(hateCrimeSubset)) {if(hateCrimeSubset$Pop3[i]>=1000000){
  hateCrimeSubset$City.size[i]<- "Metropolitan Areas"}
}
```

**Research Question 1**
Can Region, City Size, or the interaction between Region and City Size predict hate crimes for each category of hate crime? 

Race
```{r}
lmRace<- lm(Race~ Region + City.size + Region:City.size, data=hateCrimeSubset)
summary(lmRace)
#P-value is extremely small at p-value: < 2.2e-16. We reject the null hypothesis. 
```

```{r}
library(lsr)
etaSquared(lmRace)
#When we look at eta squared, we find that City.size has a large effect and the interaction between Region and City.size has a large (though smaller) effect. Region alone does not have a small effect. 
```


```{r}
library(ggeffects)
predict1<- ggpredict(lmRace, terms=c("Region", "City.size"))
plot(predict1)
```


Religion
```{r}
lmReligion<- lm(Religion~ Region + City.size + Region:City.size, data=hateCrimeSubset)
summary(lmReligion)
#P-value is extremely small at p-value: < 2.2e-16. We reject the null hypothesis.
```

```{r}
library(lsr)
etaSquared(lmReligion)
#We see that Region has a small effect, City.size has a large effect, and the interaction between Region and City.size also have a large effect. 
```

```{r}
predict2<- ggpredict(lmReligion, terms=c("Region", "City.size"))
plot(predict2)
```

Sexual Orientation
```{r}
lmSO<- lm(Sexual.orientation~ Region + City.size + Region:City.size, data=hateCrimeSubset)
summary(lmSO)
#P-value is extremely small at p-value: < 2.2e-16. We reject the null hypothesis.
```

```{r}
library(lsr)
etaSquared(lmSO)
#Region has a small effect; City.size and the interaction between Region and City.size have large effects. 
```

```{r}
predict3<- ggpredict(lmSO, terms=c("Region", "City.size"))
plot(predict3) + labs(x="Region", y="Sexual Orientation", title = "Predicted Values of Sexual Orientation", subtitle = "Linear Model")
```

Ethnicity
```{r}
lmEthn<- lm(Ethnicity~ Region + City.size + Region:City.size, data=hateCrimeSubset)
summary(lmEthn) 
#P-value is extremely small at p-value: < 2.2e-16. We reject the null hypothesis.
```

```{r}
library(lsr)
etaSquared(lmEthn)
#Region does not have an effect; City.size and the interaction between Region and City.size have large effects.
```

```{r}
predict4<- ggpredict(lmEthn, terms=c("Region", "City.size"))
plot(predict4)
```

Disability
```{r}
lmDisa<- lm(Disability~ Region + City.size + Region:City.size, data=hateCrimeSubset)
summary(lmDisa)
#P-value is extremely small at p-value: < 2.2e-16. We reject the null hypothesis.
```

```{r}
library(lsr)
etaSquared(lmDisa)
#Region does not have an effect; City.size has a small effect, and the interaction has a medium effect. 
```

```{r}
predict5<- ggpredict(lmDisa, terms=c("Region", "City.size"))
plot(predict5)
```

Gender
```{r}
lmGen<- lm(Gender~ Region + City.size + Region:City.size, data=hateCrimeSubset)
summary(lmGen)
#p-value is extremely small at 4.341e-12; we reject the null hypothesis.
```

```{r}
library(lsr)
etaSquared(lmGen)
#Region does not have an effect. City.size has a small effect. The interaction has a medium effect. 
```

```{r}
predict6<- ggpredict(lmGen, terms=c("Region", "City.size"))
plot(predict6)
```

Gender Identity
```{r}
lmGI<- lm(Gender.Identity~ Region + City.size + Region:City.size, data=hateCrimeSubset)
summary(lmGI)
#P-value is extremely small at p-value: < 2.2e-16. We reject the null hypothesis.
```

```{r}
library(lsr)
etaSquared(lmGI)
#Region does not have an effect. City.size has a small effect. The interaction has a large effect. 
```

```{r}
predict7<- ggpredict(lmGI, terms=c("Region", "City.size"))
plot(predict7) + labs(x="Region", y="Gender Identity", title = "Predicted values of Gender Identity", subtitle = "Linear Model")
```

**Research Question 2**
Do the quarters have a statistically significant difference in mean number of overall hate crimes?  

```{r}
longHate <- hateCrimeSubset %>%
  pivot_longer(cols = Race:Gender.Identity,   
               names_to = "HateCrimeType",  
               values_to = "HateCrimeValue") 

longHate <- longHate %>%
  pivot_longer(cols = starts_with("X"),  
               names_to = "Quarter",    
               values_to = "HateCrimeCount") 
```


```{r}
avR2<- aov(HateCrimeCount~ Quarter, longHate)
summary(avR2)
```

```{r}
TukeyHSD(avR2)
#Statistically significant difference between 2nd and 1st quarter; 3rd and 1st; 4th and 2nd; 4th and 3rd. This difference could be the result of factors which it would be necessary for further research to analyze for potential contributions to preventive measures hate crimes. 

```

```{r}
mean(hateCrimeSubset$X1st.quarter, na.rm =TRUE)
mean(hateCrimeSubset$X2nd.quarter, na.rm =TRUE)
mean(hateCrimeSubset$X3rd.quarter, na.rm =TRUE)
mean(hateCrimeSubset$X4th.quarter, na.rm =TRUE)
```

```{r}
longHateDetails <- longHate %>% group_by(Quarter) %>% summarise(hateCrimeMean = mean(HateCrimeCount, na.rm = TRUE), SE = sd(HateCrimeCount, na.rm = TRUE) / sqrt(n()))
```


```{r}
longHateDetails$Quarter<- as.factor(longHateDetails$Quarter)
levels(longHateDetails$Quarter)<- c("First Quarter", "Second Quarter", "Third Quarter", "Fourth Quarter")
```

```{r}
library(ggplot2)
```

```{r}
ggplot(longHateDetails, aes(Quarter, hateCrimeMean)) + geom_bar(stat="identity", position = "dodge", color="sienna", fill="burlywood") + theme_linedraw() + labs(y="Overall Quarter Hate Crime Count", x="Quarter", title = "Total Hate Crime Count Across Quarters", subtitle = "Anova") +geom_errorbar(aes(ymin=hateCrimeMean- SE, ymax = hateCrimeMean+ SE, width = 0.2))
```


**Research Question 3**
Is there a positive relationship between city population and overall hate crime across quarters? 

Analyses: 
```{r}
library(ltm)
cor.test(hateCrimeSubset$X1st.quarter, hateCrimeSubset$Pop3, method = "spearman")
#Rho is 0.287
#P-value is extremely small at p-value: < 2.2e-16. We reject the null hypothesis.
```

```{r}
cor.test(hateCrimeSubset$X2nd.quarter, hateCrimeSubset$Pop3, method = "spearman")
#0.287
#P-value is extremely small at p-value: < 2.2e-16. We reject the null hypothesis.
```

```{r}
cor.test(hateCrimeSubset$X3rd.quarter, hateCrimeSubset$Pop3, method = "spearman")
#0.267
#P-value is extremely small at p-value: < 2.2e-16. We reject the null hypothesis.
```

```{r}
cor.test(hateCrimeSubset$X4th.quarter, hateCrimeSubset$Pop3, method = "spearman")
#0.283
#P-value is extremely small at p-value: < 2.2e-16. We reject the null hypothesis.
```


As we can see, there is a positive correlation. 
```{r}
longHate$Quarter<-as.factor(longHate$Quarter)
levels(longHate$Quarter)<-c("1st Quarter", "2nd Quarter", "3rd Quarter", "Fourth Quarter")
levels(longHate$Quarter)
```

```{r}
library(ggplot2)
ggplot(longHate, aes(HateCrimeCount, Pop3)) + geom_point() + theme()  + labs(y="City Population", x="Hate Crime Count", title = "Hate Crime Count by City Population Size", subtitle="Spearman's Rho") +  geom_smooth(method = "lm", color = "red", linetype = "dashed") + facet_wrap(~Quarter)
```

